---
title: "Quality Control Example"
author: "Robert M Flight"
date: "`r Sys.time()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quality Control Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{visualizationQualityControl, ggplot2, ggbiplot, ComplexHeatmap}
---

## Introduction

This vignette shows how the various functions in this package can be used for 
basic quality control, in addition to advocating a particular workflow for 
examining experimental data prior to analysis. These steps consist of:

* Principal components analysis
* Correlation heatmap
* Median correlation
* Feature outliers

In all of the examples, we will compare and contrast a dataset that is composed
of two different conditions, and the same dataset wherein two samples have had
their class labels switched by mistake, and how the visualizations above can
illustrate potential problems.

## Data

```{r load_data}
library(visualizationQualityControl)
data(grp_cor_data)
exp_data <- grp_cor_data$data
rownames(exp_data) <- paste0("f", seq(1, nrow(exp_data)))
colnames(exp_data) <- paste0("s", seq(1, ncol(exp_data)))

sample_info <- data.frame(id = colnames(exp_data), class = grp_cor_data$class)
sample_classes <- sample_info$class

mix_data <- exp_data
mix_data[, 5] <- exp_data[, 19]
mix_data[, 19] <- exp_data[, 5]
```


The data used here are from a simulated experiment, where there are two classes
of samples. `r nrow(exp_data)` features measured, across `r ncol(exp_data)` samples
from `r length(unique(sample_classes))` groups, with `r sum(sample_classes == "grp1")`
in each group.

The actual values we are going to use are in `exp_data`. The classes of samples are
defined in `sample_classes`. We will also create a version of the data that has
two samples completely switched by accident, labels and all, in `mix_data`.

```{r show_data}
str(exp_data)
exp_data[1:5, 1:5]
sample_info
```

